A Switching Planner for Combined Task and Observation Planning

### 2010-11-28 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
Few fixes requested by Moritz

Planning for agents acting in partially observable environments with
stochastic actions is extremely computationally
expensive \cite{mdp-complexity}. Even relatively small partially
observable Markov decision problems (POMDPs), represented by tens of
propositional variables, produce state spaces far larger than existing
optimal or close to optimal planners can handle. For the easier
problem of planning in completely observable stochastic domains, an
alternative approach has been to use classical sequential planners and
replan when actions have unplanned-for
outcomes \cite{yoon:etal:2007}. However, these approaches are not as
useful in partially observable domains as replanning is triggered when
the system ends up in a state other than the desired one, and this
state cannot in general be determined.

In this paper we describe, \pcogx, the planning component in a robotic
system that continuously deliberates in a stochastic dynamic
environment in order to achieve objectives set by the user, and
acquire knowledge about its surroundings. This domain has the
characteristics of a POMDP in that the state is not perfectly known,
and state information is gained by the robot through the use of
unreliable sensing actions. However, the state representation is
propositional, and even for relatively simple problems is too large
for a POMDP solver to be applied.

Inspired by the replanning approaches used successfully for MDP
planning \cite{yoon:etal:2007,yoon:etal:2008} we propose an approach
that uses a replanning sequential planner to build the majority of the
plan, but uses a decision-theoretic planner on small subproblems where
reasoning about observations is required to find good plans. We refer
to the complete system as a {\em switching planner}. The approach is
not optimal, particularly as it is based on a satisficing sequential
planner, but it performs better than the sequential planner alone and
is fast enough to be used on the robot for decision-making. As an
illustration, consider a robot searching for a box that it knows is
often found in offices. The sequential planner might produce a plan
that assumes that {\sc Room a} is an office, that the box is in {\sc
Room a}, and that the box is in location 1 in {\sc Room a}, and then
goes to {\sc Room a}, goes to location 1 and searches for the
box. When the plan gets executed and the search begins, the
decision-theoretic planner is called to conduct the search, and at
this point may retract the last assumption and build a plan to search
the whole of {\sc Room a} for the box. If this plan fails, it will do
so by disproving one of the assumptions (for example, that {\sc Room
a} is an office), which will then lead to replanning by the sequential
planner based on the newly discovered knowledge.

To represent planning problems of this kind, we have developed a
first-order declarative language, called decision-theoretic planning
domain definition language (DTPDDL), for describing domains of
planning problems that correspond to POMDPs. We have created a system
that automatically constructs from this a determinised planning domain
for the sequential planner that includes actions that allow
assumptions to be made about the values of imperfectly known state
variables. These are then used to restrict the state space and set the
reward function for the decision-theoretic planner, which can operate
with the original representation of the domain.

The sequential planner we use is Continual
Planning \cite{brenner:nebel:jaamas09} (CP). CP is based
on \fastdownward \cite{fast-downward}. CP already has the capability
to replan, and also allows agent knowledge to be represented
explicitly, so we can write actions that gain the agent knowledge. For
the decision-theoretic planner we use an information-state
\laostar\ procedure that uses the solution to the equivalent MDP as a
heuristic. For reasons of space, we will not describe this system in
detail here.

The switching planner approach gives us some important advantages in
our robotic domain. First, the main source of uncertainty in the
domain is unreliable observations, rather than stochasticity in the
action effects, and this approach is particularly suited to problems
that combine a deterministic task planning problem with observations
as the decision-theoretic planner is only invoked to determine the
values of variables. Secondly, the replanning makes the system very
robust to changing goals and the discovery of new facts about the
world, both of which are features of our domain.

@@

An outline of the paper is as follows. We begin by describing POMDPs with a
propositionally factored representation of states and observations,
and the relationship between this and a flat POMDP representation. Next
we describe our DTPDDL language using a running example of the robot
looking for a box. Then we describe the switching planner, including
how the sequential domain is produced and a plan is generated, and how
this then provides the input for the decision-theoretic
planner. Finally we provide an evaluation of the system on the example
robot domain, and finish with a short survey of related work and some
concluding remarks.

>>>>

Planning for agents acting in partially observable environments with
stochastic actions is extremely computationally
expensive \cite{mdp-complexity}. Even relatively small partially
observable Markov decision problems (POMDPs), represented by tens of
propositional variables, produce state spaces far larger than existing
optimal or close to optimal planners can handle. For the easier
problem of planning in completely observable stochastic domains, an
alternative approach has been to use classical sequential planners and
replan when actions have unplanned-for
outcomes \cite{yoon:etal:2007}. However, these approaches rely
crucially on being able to determine properties of the current state
in order to decide to replan. For partially observable domains, only
probabilistic information about these properties will in general be
available.

In this paper we describe, \pcogx, the planning component in a robotic
system that continuously deliberates in a stochastic dynamic
environment in order to achieve objectives set by the user, and
acquire knowledge about its surroundings. This domain has the
characteristics of a POMDP in that the state is not perfectly known,
and state information is gained by the robot through the use of
unreliable sensing actions. However, the state representation is
propositional, and even for relatively simple problems the underlying
model is too large for a POMDP solver to be applied.

Inspired by the replanning approaches used successfully for MDP
planning \cite{yoon:etal:2007,yoon:etal:2008} we propose an approach
that uses a replanning sequential planner to build the majority of the
plan, but uses a decision-theoretic planner on small subproblems where
reasoning about observations is required to find good plans. We refer
to the complete system as a {\em switching continual planner}. The
approach is not optimal, particularly as it is based on a satisficing
sequential planner, but it performs better than the sequential planner
alone and is fast enough to be used in real-time on a mobile robot for
decision-making.

As an illustration, consider a robot searching for a box that it knows
is often found in offices. The sequential planner might produce a plan
that assumes that {\sc Room a} is an office, that the box is in {\sc
Room a}, and that the box is in location 1 in {\sc Room a}. Given
these assumptions, the plan might be to go to {\sc Room a}, go to
location 1 and search for the box. When the plan gets executed and the
search begins the decision-theoretic planner is called since searching
involves making observations. The decision-theoretic planner might
retract the last assumption but leave the others, and build a plan to
search the whole of {\sc Room a} for the box. If this plan fails to
find the box, it will disprove the assumption that the box is in {\sc
Room a}, which will then lead to replanning by the sequential planner
based on the newly discovered knowledge.

To represent planning problems of this kind, we have developed a
first-order declarative language, called decision-theoretic planning
domain definition language (DTPDDL), for describing domains of
planning problems that correspond to POMDPs. We have created a system
that automatically constructs from this a deterministic planning domain
for the sequential planner that includes actions that allow
assumptions to be made about the values of imperfectly known state
variables. These are then used to restrict the state space and set the
reward function for the decision-theoretic planner, which can operate
with the original representation of the domain.

The sequential planner we use is based
on \fastdownward \cite{fast-downward}. We add the capability to
replan, and also allow agent knowledge to be represented explicitly,
so we can write actions that gain the agent knowledge. For the
decision-theoretic planner we use forward search in the information
state.

The switching planner approach gives us some important advantages in
our robotic domain. First, the significant source of uncertainty in
the domain is the unreliability of observations, and this approach is
particularly suited to problems that combine a deterministic task
planning problem with observations as the decision-theoretic planner
is only invoked to determine a characteristic of the underlying
state. Secondly, the replanning makes the system very robust to
changing objectives and the discovery of new facts about the world, both of
which are features of our domain.

The remainder of the paper proceeds as follows. We begin by describing
POMDPs with a propositionally factored representation of states and
observations. We also outline the relationship between this and a flat
POMDP representation. Next we describe our DTPDDL language with
example declarations from a mobile robot exploration task. Then we
describe the switching planner, including how the sequential domain is
produced and a plan is generated, and how this then provides the input
for the decision-theoretic planner. Finally we provide an evaluation
of the system on the example robot domain, and finish with a short
survey of related work and some concluding remarks.

@@

### 2010-11-28 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Few fixes requested by Moritz

In a second step, we iteratively refine the relaxed declaration by
adding terms from the original statement of $(\pp{:init})$ while the
number of abstract states that occur with non-zero probability
according to that refined declaration remains of a practicable
size. In detail, for each excluded atomic term, we compute the {\em
entropy} of the corresponding proposition, {\em conditional} on the
active assumptions. The {\em conditional entropy} measure
is higher the more uncertain that proposition is given truth
assignments to propositions that are the subject of active
assumptions. Propositions that in the first place were excluded are
iteratively added to the problem posed to the conditional session in
increasing order according to that measure.

>>>>

In a second step, we iteratively refine the relaxed declaration by
adding terms from the original statement of $(\pp{:init})$ while the
number of abstract states that occur with non-zero probability
according to that refined declaration remains of a practicable
size. In detail, for each excluded atomic term, we compute the {\em
entropy} of the active assumptions, {\em conditional} on the
corresponding proposition. Loosely speaking, the {\em conditional
entropy} measure is higher the more uncertain those assumptions are
given truth assignment to the proposition. Propositions that in the
first place were excluded are iteratively added to the problem posed
to the conditional session in increasing order according to that
measure.

%% In detail, for each excluded atomic term, we compute the {\em
%% entropy} of the corresponding proposition, {\em conditional} on the
%% active assumptions. The {\em conditional entropy} measure
%% is higher the more uncertain that proposition is given truth
%% assignments to propositions that are the subject of active
%% assumptions. Propositions that in the first place were excluded are
%% iteratively added to the problem posed to the conditional session in
%% increasing order according to that measure.
